---
title: "Tooth Growth Basic Inferential Data Analysis"
author: "Kaushal Yadav"
date: "13 December 2018"
output: html_document
---

##Overview
The purpose of the this data analysis is to analyze the ToothGrowth data set by
comparing the guinea tooth growth by supplement and dose. First, I will do
exploratory data analysis on the data set. Then I will do the comparison with
confidence intervals in order to make conclusions about the tooth growth.

##Instruction
Analyze the ToothGrowth data in the R datasets package. Tasks

1- Load the ToothGrowth data and perform some basic exploratory data analyses
2- Provide a basic summary of the data.
3- Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
4- State your conclusions and the assumptions needed for your conclusions.

##Set-Up
Load required packages.
Load required data: Dataset - The Effect of Vitamin C on Tooth Growth in Guinea Pigs.
```{r}
ECHO=TRUE
library(ggplot2)
library(datasets)
data("ToothGrowth")
```

###Exploring the Data
Description The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).
```{r}
str(ToothGrowth)
head(ToothGrowth)
```
The data format can be improved on - and especially this will help us with the plotting: The dose variable has three levels. Therefore it makes more sense to re-format it to a factor, than keep in nummerical.
So let’s do that, summarize the data and make a quick plot (please note that the x-axis is now not gradually increasing, but is showing the three levels)
```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
summary(ToothGrowth)
```

```{r}
qplot(dose ,len ,data = ToothGrowth, 
      col = supp, 
      main = "Tooth growth of guinea pigs by supplement type and dosage (mg)", 
      xlab = "Dosage (mg)", 
      ylab = "Tooth length")
```
To see the difference better, let’s put the same data in a box-plot-style graph. (remark: to have the facets, it was a good idea to re-format the dose variable)
```{r}
qplot(supp, len, data = ToothGrowth, 
      facets = ~dose, 
      main = "Tooth growth of guinea pigs by supplement type and dosage (mg)", 
      xlab = "Supplement type", 
      ylab = "Tooth length") + 
        geom_boxplot(aes(fill = supp))
```
1. we see that the OJ generally is doing better in each dosage compared to VC
2. increasing the dosage (from 0.5, to 1, to 2) increased the length of the tooth, for both supplement types.

####Comparing tooth growth by supp and dose
##Hypothesis 1
Let H0: No difference in tooth growth given OJ or VC
Ha: tooth growth is bigger when using OJ, than VC.
```{r}
VC.length <- ToothGrowth$len[ToothGrowth$supp == "VC"]
OJ.length <- ToothGrowth$len[ToothGrowth$supp == "OJ"]
```
####Student's t-Test
```{r}
t.test(OJ.length, VC.length, 
       alternative = "greater",
       paired = FALSE,
       var.equal = FALSE,
       conf.level = 0.95)
```
The p-values of this comparison is p = 3%, lower than 5%. We reject the null hypothesis.
So we conclude that the alternative hypothesis is true: OJ has a greater impact on tooth growth than VC.

##Hypothesis 2
Let H0: No difference in growth given different doses
Ha:Different growth given different doses
```{r}
dose_0.5 <- ToothGrowth$len[ToothGrowth$dose == "0.5"]
dose_1   <- ToothGrowth$len[ToothGrowth$dose == "1"]
dose_2   <- ToothGrowth$len[ToothGrowth$dose == "2"]
```
Step 1: perform a t-test between dose_0.5 and dose_1
```{r}
t.test(dose_0.5, dose_1, 
       alternative = "less",
       paired = FALSE,
       var.equal = FALSE,
       conf.level = 0.95)
```
The p-value is very small (6.342e-8), therefore we can conclude that the null hypothesis can be rejected looking at dose_0.5 and dose_1.

Step 2: perform a t-test between dose_1 and dose_2:
```{r}
t.test(dose_1, dose_2, 
       alternative = "less",
       paired = FALSE,
       var.equal = FALSE,
       conf.level = 0.95)
```
Here too, the p-value is very small, therefore we can safely reject the null hypothesis.
We can conclude that the higher the dosage gets, the bigger the tooth growth gets. Hence, we reject the null Hypothesis and accept the alternative hypothesis.

####Conclusion
We can summarize that
- there is (at least) a 95% confidence that by increasing the dosage from 0.5 to 1mg and from 1 to 2mg, increases the tooth length.
- there is (at least) a 95% confidence that giving the supplement OJ (Orange Juice) increases the tooth length more significant than giving VC (Vitamin C)


